### 2020-11-09 police_modeling.Rmd
results

---
title: "Modeling"
author: "Connie Wu"
date: "10/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Modeling
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(lmtest)
library(lme4)
library(caret)
library(groupdata2) # fold()
```

```{r load-data}
final_df <- readRDS("data/final.rds")

# Combining Asian and Other so model will converge
final_df$SubjectRace[final_df$SubjectRace == "A"] <- "O"
final_df$OfficerRace[final_df$OfficerRace == "A"] <- "O"

final_df <- final_df %>% 
  mutate(so = as.numeric(str_count(final_df$SubjectRace, "O|A") > 0),
         oo = as.numeric(str_count(final_df$OfficerRace, "O|A") > 0)) %>% 
  select(-c(sa, oa))

single_victims <- final_df %>% 
  filter(NumberOfSubjects <= 1) %>% 
  mutate(SubjectRace = relevel(as.factor(SubjectRace), "W"))

mult_victims <- final_df %>% 
  filter(NumberOfSubjects > 1)
```

### Old Validation Code

```{r}
crossvalidate <- function(data, k, model, dependent, random = FALSE){
  # 'data' is the training set with the ".folds" column
  # 'k' is the number of folds we have
  # 'model' is a string describing a linear regression model formula
  # 'dependent' is a string with the name of the score column we want to predict
  # 'random' is a logical; do we have random effects in the model?

  # Initialize empty list for recording performances
  performances <- c()

  # One iteration per fold
  for (fold in 1:k){

    # Create training set for this iteration
    # Subset all the datapoints where .folds does not match the current fold
    training_set <- data[data$.folds != fold,]

    # Create test set for this iteration
    # Subset all the datapoints where .folds matches the current fold
    testing_set <- data[data$.folds == fold,]

    ## Train model

    # If there is a random effect,
    # use glmer() to train model
    # else use glm()

    if (isTRUE(random)){
      # Train linear mixed effects model on training set
      m <- glmer(model, training_set, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
    } else {
      # Train linear model on training set
      m <- glm(model, training_set, family = "binomial", control=glm.control(maxit=50))
    }

    ## Test model

    # Predict the dependent variable in the testing_set with the trained model
    predicted <- predict(m, testing_set, type="response", allow.new.levels = TRUE)
    fit.convert <- ifelse(predicted < 0.5, 0, 1)
    table(fit.convert, testing_set[[dependent]])
    performances[fold] <- mean(fit.convert == testing_set[[dependent]])
  }

  return(performances)
}
```

```{r}
getAccuracy <- function(v) {
  if (length(v) == 4) {
      return(((v[1] + v[4])/sum(v)) * 100)
  }
  if (length(v) == 2) {
    return((v[1]/sum(v)) * 100)
  }
}
```

## Single Victims

### Backwards Selection
```{r backwards-selection}
fullmodel <- glm(fatal ~ SubjectRace + SubjectGender + ob + ow + ol + oo + omale + ofem + omult + gun + weapon_other + replica + knife + unarmed + I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train, family = binomial, data=single_victims)
step(fullmodel, direction = "backward")
```

Backwards selection gives: glm(formula = fatal ~ SubjectRace + SubjectGender + ob + ow + 
    omult + gun + knife + unarmed + I(black/city_pop_2013), family = binomial, 
    data = single_victims)

```{r backwards-selection}
fullmodel <- glm(fatal ~ SubjectRace + SubjectGender + ob + ow + ol + oo + omale + ofem + omult + gun + weapon_other + replica + knife + unarmed + I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train + city, family = binomial, data=single_victims)
step(fullmodel, direction = "backward")
```

When I put city in, it turns out to be in the backwards model too -- should probably do random effects with city. Formula: glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city, family = binomial, data = single_victims)

### Trying the models

```{r backwards-fit}
fitback <- glm(formula = fatal ~ SubjectRace + SubjectGender + ob + ow + 
    omult + gun + knife + unarmed + I(black/city_pop_2013), family = binomial, 
    data = single_victims)
summary(fitback)
```

```{r validate fitback}
set.seed(1)
Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

mod_fitback <- train(as.factor(fatal) ~ SubjectRace + SubjectGender + ob + ow + 
    omult + gun + knife + unarmed + I(black/city_pop_2013),  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
    )

pred <- predict(mod_fitback, newdata=testing )
confusionMatrix(data=pred, as.factor(testing$fatal))
```

```{r validate fitback again}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + SubjectGender + ob + ow + omult + gun + knife + unarmed + I(black/city_pop_2013)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = FALSE)
p0
mean(p0)

back2 <- glm(m0, training, family = "binomial")
preds <- predict(rand_fit1, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r fit with city}
fitback1 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city, family = binomial, data = single_victims)
summary(fitback1)

back2 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city + city:SubjectRace, family = binomial, data = single_victims)
lrtest(fitback1, back2)

back3 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city + city:gun, family = binomial, data = single_victims)
lrtest(fitback1, back3)
```

Interactions between city and race lead to better fit, makes it seem like I should vary race by city perhaps

```{r validate fitback1}
set.seed(1)
Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

mod_fitback1 <- train(as.factor(fatal) ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city,  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
    )

pred <- predict(mod_fitback1, newdata=testing )
confusionMatrix(data=pred, as.factor(testing$fatal))
```
```{r validate back2}
set.seed(1)
Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

mod_back2 <- train(as.factor(fatal) ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city + city:SubjectRace,  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
    )

pred <- predict(mod_back2, newdata=testing )
confusionMatrix(data=pred, as.factor(testing$fatal))
```

```{r random-effects}
rand_fit1 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + (1 | city), data = single_victims, family = "binomial", nAGQ=50)
summary(rand_fit1)
coef(rand_fit1)
fixef(rand_fit1)
ranef(rand_fit1)

classDFrand_fit1 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit1),0))
rand_fit1pred <- xtabs(~ predicted + response, data = classDFrand_fit1)
getAccuracy(rand_fit1pred)
```

```{r validate rand_fit1}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 | city)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit1 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit1, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r vary subject race by city}
rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(rand_fit2)
coef(rand_fit2)
fixef(rand_fit2)
ranef(rand_fit2)

classDFrand_fit2 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit2),0))
rand_fit2pred <- xtabs(~ predicted + response, data = classDFrand_fit2)
getAccuracy(rand_fit2pred)
```

```{r validate rand_fit2}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (SubjectRace | city)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit2 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit2, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r vary subject race by city and make it vary intercept too}
rand_fit3 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + (1 + SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(rand_fit3)
coef(rand_fit3)
fixef(rand_fit3)
ranef(rand_fit3)

classDFrand_fit3 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit3),0))
rand_fit3pred <- xtabs(~ predicted + response, data = classDFrand_fit3)
getAccuracy(rand_fit3pred)
```

```{r validate rand_fit3}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + SubjectRace | city)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit3 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit3, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r random effects 4}
rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(rand_fit4)
coef(rand_fit4)
fixef(rand_fit4)
ranef(rand_fit4)

classDFrand_fit4 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit4),0))
rand_fit4pred <- xtabs(~ predicted + response, data = classDFrand_fit4)
getAccuracy(rand_fit4pred)
```

```{r validate rand_fit4}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit3 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit3, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

Single victims final model!!

## Multiple Victims
```{r backwards-selection}
fullmodel <- glm(fatal ~ sb  + sw + sl + so + sfem + smale + weapon_other + replica + knife + ob + ow + ol + oo + omale + ofem + omult + gun + unarmed + unknown_weapon +  I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train, family = binomial, data=mult_victims)
step(fullmodel, direction = "backward")
```